Findings

### 2019-02-22 NB_2r2c.Rmd
update

---
title: "Naive Bayes Analysis Ver 2r.2c"
author: "Dan Johnson"
date: "2/14/2019"
output: html_document
---

@@

##Summary
This workflow uses Naive Bayes to classify TA comments based on Subject and Structure. The major additions and revisions in **Version 2r.2c** are:

*  Randomly selected data rows used for training and test datasets. 
*  Final output matrix tables are more informative. 
*  Addition of a back-analysis of entire dataset, not just testing sub-set.
*  Final output CSV file is a tidy data table of the original comments, the hand-scored and predicted subject, and hand-scored and predicted structure.

@@

```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)

#library(RTextTools)
#library(doMC)
#registerDoMC(core=detectCores())
#library(tidyr)
#library(dplyr)
#library(gmodels)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r data_input}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data as subsets of CSV.
```{r data_subsetting}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

@@

```{}
# Function to KEEP word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead.
convert_counts2 <- function(x) {
  y <- x
}
```

```{r make_binary}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r bayes_training}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

@@

#The truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject. this is what the confusion matrix calculates. The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.

#The Cross Table command below (from "gmodels") generates a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

@@

##Graph Outputs
Plotting Hand-coding vs. Bayes

Need to combine two datasets into one, and create subsets based on the

```{r}
tidy(proptallysum1)
ggplot(data=proptallysum1, aes(x=Hand.Coded.Subject, y=total, group=1))+
  geom_line() +
  geom_point()

ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
  geom_line() +
  geom_point()
 ```

>>>>

---
title: "Naive Bayes Analysis Ver 2r.2c"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

@@

##Summary
This it the current Naive Bayes workflow used to classify TA comments based on Subject and Structure. Additions and revisions made to **Version 2r.2c** are:

*  Random selection of data rows used for training and test datasets. 
*  Final output matrix tables are more informative. 
*  Addition of a back-analysis step that reviews entire dataset, not just testing sub-set.
*  Outputs a tidy data table in CSV format that contains the original comments (pre-clean-up) plus their hand-scored and predicted subject and structure.

@@

```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r input data}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data as subsets of CSV.
```{r subset data}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

@@

```{}
# Alternate version of convert_function code block. This version KEEPs word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead. 
#Only works when the NB classifier is looking at ingroup/outgroup. Breaks on multiple categories.
#
convert_counts2 <- function(x) {
  y <- x
}
```

```{r apply_conversion}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

@@

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r bayes_training}
# Train the classifier on training data, then test on test subset and full dataset.
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

@@

#The truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject. this is what the confusion matrix calculates.

#The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.

#Alternatively, the Cross Table command below (from "gmodels" package) generates a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

@@

##Graph Outputs
Comparing outcome from Hand-coding vs. Bayes

Need to combine two datasets into one.

```{r}
ggplot(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`,group=1))+
  geom_line()
ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
  geom_line()
ggplot(data=proptallysum3, aes(x=Hand.Coded.Structure, y=total, group=1))+
  geom_line()
ggplot(data=proptallysum4, aes(x=Bayes.Predicted.Structure, y=total, group=1))+
  geom_line()
```

### 2019-02-22 findings.Rmd
update

---
title: "Findings"
author: "Dan Johnson"
date: "2/22/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

##Classifier Accuracy
The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening. The far right columns are overall subject frequencies, the bottom rows, overall structure frequencies.

@@

>>>>

---
title: "Findings"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

#Classifier Accuracy
*Updated 2/22/19*

The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening. The far right columns are overall subject frequencies, the bottom rows, overall structure frequencies.

@@

#Overview
This notebook summarizes my progress, thoughts while building a working Ensemble method for categorizing TA comments.

##Replicate Cycles Test of Naive Bayes
One of my ideas was to see if I could refine Naive Bayes by running the same classifier with different randomized datasets, then putting individual outputs side by side and assigning final categories based on most frequent classification in the set.

To test this I ran 5 Subject classifier replicates using the code in *Naive_bayes_V2r_2b.rmd*, with 5 different ```set.seed``` values. The exported CSV files were combined to get a set of data with up to 5 classifier assignments for the same comment.

Using multiple randomized NB training runs does not seem to help. First, it was rare for the  randomized runs to assign different incorrect classes. Second, when I compared 5 independently randomized runs by eye, the incorrectly categorized items were repeatedly put into the SAME incorrect group each time. Based on this, an Ensemble method using H-stacked categorical NB is unlikely to reduce overall errors in classification.

Looking at the combined data DID provide other helpful insights into the overall classification procedure though. Naive Bayes does not seem to do well with comments that:

*  Contain many or mostly punctuation symbols, such as:
    +  : not ,
    +  ?
    +  ??
    +  ???
*  Are whole numbers
    +  1
    +  2
    +  10
*  Are one word questions or comments
    +  because?
    +  "and"
*	 Are sentences enclosed in punctuation
*  Are made up entirely of STOP/common words (I am not as sure on this one)

##Overall Accuracy
I also looked at OVERALL accuracy of NB by recreating the 2-dimensional contingency table that I made when hand-coding, and comparing it side-by-side to NB classifications. From this I gained additional insights.

OVERALL, the 2-D Naive Bayes model performed very well for classifying comment Subject. When I re-ran the full original dataset (9340 comments) through the prediction model, NB estimates for Subject were within 1.5% of my hand-scored assignments. Structure classification was not as robust. Other observations:

*  NB over-classified comments as "copyediting of technical issues" by 8-10%, and under-classified comments that should be marked as "specific technical issues" by almost same amount.
*  Comments that belonged to groups making up less than 1-2% of all comments were more likely to be mis-classified. 
*  Longer text comments seemed to be mis-classified more often. Follow-up methods to evaluate: 
    +  Split longer comments into shorter phrases
    +  Evaluate only the first 5-10 words of a comment

It appears that Subject categories are more terms-centered, while Structure is based more on the sentiment and sentence structure of a comment. This suggests that I need to incorporate a sentiment classifier or PoS tagger step.

##Mixed Ensemble Method
Based on the  seems like the most reasonable approach for maximum accuracy is going to be a mixed Ensemble method. The initial working model is to:

Add a pre-screening step prior to implementing NB:

*  Isolate "short-string" comments first. These are comments consisting of single words, two-word phrases, more punctuation than text, etc. These represent mostly pointers and copy-editing.
*  Look for 2-grams with high frequencies in one particular category. (Ex., "basic criteria" 2-gram identifies a Basic Criteria comment.)
